---
title: "Class Assignment: MPG tables and charts (edited with dplyr functions)"
author: "Jacob Burke"
date: "September 20th, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)                ## loading the entire knitr package
library(tidyverse)              ####### change# load tidyverse
library(esquisse)             ## use esquisse to setup the basic plots

library(kableExtra)
library(magrittr)
opts_chunk$set(echo = TRUE)
```

# Using the tidyverse (especially dplyr) for MPG

Recall assignment:

Tables should include a comparison of city and highway mileage by the class of car and the best three car models for city and highway mileage for all the years in which data is available.  

Plot the data displaying as much of the data as you can. Put continuous variables on the axes and include a locally smoothed regression line to show the relationship with mileage.  Also make a box plot comparing city and highway MPG by class of car.  


## Tables

```{r}

data(mpg)     ## load the data

mpg$cyl <- as.factor(mpg$cyl)   ## convert discrete variables to factors so that they plot
mpg$drv <- as.factor(mpg$drv)   ## as different colors, not gradations of one color

mpg$cty <- as.double(mpg$cty)

mpg_1 <- select(mpg, cty, hwy, class )


mpg_1 %<>% group_by(class) %>% 
  summarise(City=mean(cty),
            Highway=mean(hwy)) %>% 
  arrange(class, City, Highway) 


kable(mpg_1, format = "latex", booktabs=TRUE, digits = 2,     ## call kable to make the table
      col.names = c("Class", "City", "Highway"), 
      caption = "Mean City and Highway MPG by Car Class" )      

```





```{r}

## Selecting city, year, and model, variables
mpgcity <- select(mpg, model, cty, year)

## filtering down to respective year 
mpgcity_1999 <- filter(mpgcity, year == 1999)

## grouping this subset by the model, and getting each model's average Milage 
mpgcity_1999 %<>% group_by(model) %>%
              summarise(Milage = mean(cty)) %>%
              arrange(model, Milage)

## ordering data frame in order of top Milage models 
mpgcity_1999 <- mpgcity_1999[
  with(mpgcity_1999, order(-Milage)),
]

## rounding milage
mpgcity_1999$Milage <- round(mpgcity_1999$Milage, 2)


## Repeating process for mpghwy_1999
mpghwy <- select(mpg, model, hwy, year)

mpghwy_1999 <- filter(mpghwy, year == 1999)

mpghwy_1999 %<>% group_by(model) %>%
              summarise(Milage = mean(hwy)) %>%
              arrange(model, Milage)

mpghwy_1999 <- mpghwy_1999[
  with(mpghwy_1999, order(-Milage)),
]
mpghwy_1999$Milage <- round(mpghwy_1999$Milage, 2)


## ...mpgcity_2008
mpgcity_2008 <- filter(mpgcity, year == 2008)

mpgcity_2008 %<>% group_by(model) %>%
              summarise(Milage = mean(cty)) %>%
              arrange(model, Milage)

mpgcity_2008 <- mpgcity_2008[
  with(mpgcity_2008, order(-Milage)),
]
mpgcity_2008$Milage <- round(mpgcity_2008$Milage, 2)


## ...mpghwy_2008
mpghwy_2008 <- filter(mpghwy, year == 2008)

mpghwy_2008 %<>% group_by(model) %>%
              summarise(Milage = mean(hwy)) %>%
              arrange(model, Milage)

mpghwy_2008 <- mpghwy_2008[
  with(mpghwy_2008, order(-Milage)),
]
mpghwy_2008$Milage <- round(mpghwy_2008$Milage, 2)


table3 <- cbind(mpgcity_1999 [1:3, ],
               mpghwy_1999[1:3, ],
               mpgcity_2008[1:3, ],
               mpghwy_2008[1:3, ])

colnames(table3) <- c('Model', 'Milage', 
                     "Model", "Milage",
                     'Model', 'Milage',
                     "Model", "Milage"
                     )

kable(table3, digits = 2, format = "latex", booktabs=TRUE, ,caption = "Top 3 MPG Performing Cars: 1999, 2008") %>% kable_styling() %>%
  add_header_above(c("City 1999"=2, 
                     "Highway 1999"=2,
                     "City 2008"=2,
                   "Highway 2008"=2))

```




```{r}





library(ggplot2)

ggplot(mpg) +
 aes(x = displ, y = cty) +
 geom_point(aes(color=class),size=2) + geom_smooth() +
 scale_color_hue() +
 theme_minimal(base_size = 16) +
 facet_wrap(vars(year)) + 
  labs(x = "Engine Displacement",y = "City Mileage" ,title = "City MPG by Class of Car: 1999, 2008")
 
 
 
 ggplot(mpg) +
 aes(x = displ, y = hwy) +
 geom_point(aes(color=class),size=2) + geom_smooth() +
 scale_color_hue() +
 theme_minimal(base_size = 16) +
 facet_wrap(vars(year)) + 
  labs(x = "Engine Displacement", y = "Highway Mileage", title = "Highway MPG by Class of Car: 1999, 2008")
 
 

##esquisser(mpg)

```

Boxplots

```{r}



ggplot(mpg) +
 aes(x = class, y = cty) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal(base_size=16) +
  labs(x = "Car Class", y = "City Mileage", title = "City MPG by Class of Car: 1999, 2008")


ggplot(mpg) +
 aes(x = class, y = hwy) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal(base_size=16) +
  labs(x = "Car Class", y = "Highway Mileage", title = "Highway MPG by Class of Car: 1999, 2008")





```


